Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-03-26 bbbq_article.tex
Processed feedback Geert

@@

While humans only express a limited number of MHC haplotypes, and therefore any individual's immune system detects only a fraction of all possible
peptide fragments, the number of pathogenic peptides that can be detected is increased at the population level, because of the multiple and highly polymorphic MHC genes.
It is believed that this improves immunity at the population level, 
as mutations in a protein that disrupt a particular MHC presentation, 
so-called escape mutations, 
will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
% Geert: I removed the number of MHC molecules as is is more complex, 
% as the alpha chains coded by the maternal chromosome 
% of HLA-DQ, DR, DP can heterodimerize with the beta chains coded 
% by the paternal chromosome and vice versa

@@

% \paragraph{Selection of SNPs}
%
SNPs were picked based on ID number, which is linked to their discovery date. To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.[GEERT: good practice is to present the data in a chronological fashion: thus move this to figure 1 (or put in supplement)]

Per SNP, the 'protein' NCBI database was queried for the
protein sequence.
For each protein sequence, the protein topology was determined 
using PureseqTM.
Using these predicted protein topologies, the SNPs were scored to be located within or outside trans membrane domains. [GEERT: for Table 1: Tables don't have a legend, but instead have a title and footnotes below the table.]

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Elution studies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

[GEERT: I would move this subsection to below: first explain the bioinformatics predictions, and then show this as an experimental confirmation]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Prediction of TMH-derived peptide presentation in MHC-I}
[GEERT: mooier is je subheadings samen te vatten met een conclusie. Dus: TMH-derived peptides are predicted to be over-presented in MHC-I (en dan ook voor de andere subheadings)]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Prediction of TMH-derived peptide presentation in MHC-II}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evolutionary conservation}
[GEERT: zie mijn punt hierbove. Een korte conclusie als titel is veel sterker. Dus iets als: TMHs are more evolutionary conserved.]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Using a alpha value of $0.05$, as is common practice for an uninformed
expectation, we conclude that there are significantly less SNPs
in TMHs as expected by chance.
See figure \ref{fig:f_tmh_ncbi} for the percentages of TMHs in the transmembrane proteins.[GEERT: staat deze figuur hier wel goed? Niet naar boven?] The effect size is that, 
when randomly selecting SNPs,
within TMPs that are 50\% TMH (or correcting for this percentage),
per 100 SNPs found in soluble protein domains, 
one finds 91 SNPs in TMHs,
as depicted in figure \ref{fig:conservation}. [GEERT: zie mijn eerder opmerking over panels. Ik zou met panel A, B, C, etc werken.] [GEERT: hieronder had je indentieke tekst, maar verwees naar andere figuur \ref{fig:conservation_per_spanner}. Ik denk dat je deze hieronder kan samenvoegen]

@@

Although elution studies and in silico predictions both suggest that TMH-derived epitopes are presented in MHC-I and MHC-II, the molecular mechanisms of how integral membrane proteins are processed for MHC presentation are largely unknown [GEERT: cite Bianchi]. Most prominently, the fundamental principles of how TMHs are extracted from the  hydrophobic lipid environment into the aqueous vacuolar lumen, and their prior or subsequent proteolytic processing are unresolved. [GEERT: beetje kort. Suggesties? Experimental approaches? Hoeft niet super origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen. Frans kan vast helpen met mooie zin.]

@@

A final caveat in the elution studies is that there is methodological bias caused by mass spectronomy,
as hydrophobic fragments are detected less 
by mass spectronomy [GEERT: referentie, zie paper Bianchi]. This means that the number of TMH-derived epitopes
might be underestimated.

>>>>

\richel{TODO: Tables don't have a legend, but instead have a title and footnotes below the table. Modify the tables to make it so in the end}
\richel{TODO: Condense figures into bigger ones with panels, following 1 message per figure. Order these figures chronologically}

@@

While humans only express a limited number of MHC haplotypes, and therefore any individual's immune system detects only a fraction of all possible
peptide fragments, the number of pathogenic peptides that can be detected is increased at the population level, because of the multiple and highly polymorphic MHC genes.
It is believed that this improves immunity at the population level, 
as mutations in a protein that disrupt a particular MHC presentation, 
so-called escape mutations, 
will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
% Do noty give the number of MHC molecules, as is is more complex, 
% as the alpha chains coded by the maternal chromosome 
% of HLA-DQ, DR, DP can heterodimerize with the beta chains coded 
% by the paternal chromosome and vice versa

@@

% \paragraph{Selection of SNPs}
%
SNPs were picked based on ID number, which is linked to their discovery date. To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.

Per SNP, the 'protein' NCBI database was queried for the
protein sequence.
For each protein sequence, the protein topology was determined 
using PureseqTM.
Using these predicted protein topologies, the SNPs were scored to be located within or outside trans membrane domains.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMH-derived peptides are predicted to be over-presented in MHC-I}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\richel{
  First do the bioinformatics predictions,
  then the experimental confirmation 
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMH-derived peptides are predicted to be over-presented in MHC-II}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TMH-derived peptides are predicted to be presented in vivo}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\richel{
  Use this section as an experimental confirmation 
  of the bioinformatics predictions
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Human TMHs are evolutionarily conserved}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Using a alpha value of $0.05$, as is common practice for an uninformed
expectation, we conclude that there are significantly less SNPs
in TMHs as expected by chance.
See figure \ref{fig:f_tmh_ncbi} for the percentages of TMHs in the transmembrane proteins.[GEERT: staat deze figuur hier wel goed? Niet naar boven?] The effect size is that, 
when randomly selecting SNPs,
within TMPs that are 50\% TMH (or correcting for this percentage),
per 100 SNPs found in soluble protein domains, 
one finds 91 SNPs in TMHs,
as depicted in figure \ref{fig:conservation}. [GEERT: hieronder had je indentieke tekst, maar verwees naar andere figuur \ref{fig:conservation_per_spanner}. Ik denk dat je deze hieronder kan samenvoegen]

@@

Although elution studies and in silico predictions both suggest that TMH-derived epitopes are presented in MHC-I and MHC-II, the molecular mechanisms of how integral membrane proteins are processed for MHC presentation are largely unknown (\cite{bianchi2017}). Most prominently, the fundamental principles of how TMHs are extracted from the  hydrophobic lipid environment into the aqueous vacuolar lumen, and their prior or subsequent proteolytic processing are unresolved. [GEERT: beetje kort. Suggesties? Experimental approaches? Hoeft niet super origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen. Frans kan vast helpen met mooie zin.]

@@

A final caveat in the elution studies is that there is methodological bias caused by mass spectronomy,
as hydrophobic fragments are detected less 
by mass spectronomy (\cite{bianchi2017}). This means that the number of TMH-derived epitopes
might be underestimated.
